### 2024-09-26 paper/paper.qmd
updated data section + reorganized sections

---
title: "Investigating Trends of Absences in Toronto City Council Attendance between 2022 to 2026"
subtitle: "Determining the Ideal Meeting Types and Months for the 2022-2026 Term"
author: 
  - Wendy Yuan
thanks: "GitHub repo link: https://github.com/kiwindyy/CityCouncilAttendance"
date: "27 September 2024"
date-format: long
abstract: "This paper looks at trends in absences for Toronto City Council members using data from Open Data Toronto (@ ). The data was used to observe which members had the most absences, absences trends over the years, and what meeting types had the most absences. It was found that evening meetings saw the most absences and absences are predicted to see an increase over the years. Through identifying trends of absences, we can help minimize the number of absences by ___ members which is important as when City Council members are elected, they are expected to attend meetings to voice the population’s opinions."
format: pdf
toc: true
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(dplyr)
library(readr)
library(ggplot2)

cleaned_data <- read_csv(here::here("data/cleaned_data/cleaned_data.csv"))
raw_data <- read_csv(here::here("data/raw_data/unedited_data.csv"))
```

# Introduction

The Canadian parliamentary system is a democracy, where the population is represented in government by a selection of elected officials. To maintain the democratic structure, individuals should engage in political participation to have their voices represented, primarily through voting. Likewise, elected officials should uphold their duty of representing the opinions of those who voted them in. Otherwise, failing to do so will cause a decrease in trust in the government. So, it is the job of the elected members of the Toronto City Council to attend their meetings and express the concerns of the population they represent.

This paper looks at the absences of Toronto City Council members since the beginning of the 2022 term. The data used in this paper is found on Open Data Toronto (@ ). The study examines which council members have the highest rates of absence, the fluctuation of absences over time, and which meeting type has the most absences. This paper works to fill the gap of focused analysis on absences patterns within the City Council which directly relates to the effectiveness of representation by council members.

The data analysis focuses on examining patterns in the data for various causes of absences. It was found that evening meetings experienced the most absences and the trend over the years shows that absences are likely to increase in 2022. Additionally, ___ members were found to have the most absences. Rather than blaming individuals for absences, understanding these patterns allows for future targeted meeting scheduling to minimize absences. This ensures that elected officials are present to represent the voices of the general.

The paper is structured as follows: section 2 discusses the raw and clean data used in the study and any limitations in the dataset. Section 3 shows a graphical visualization of the analysis results using data from section 2. Finally, section 4 covers interpretations of each graph from section 3 and any weaknesses of this study.

# Data {#sec-data}

## Raw Data

The data used to investigate the attendance of Toronto City Council members is from the Open Data Toronto website (@ ). To measure how much City Council members worked to voice their thoughts - their attendance was examined. The attendance data is collected by sessions, with each new session beginning when a meeting is finished. A member was labelled as present if they attended any meeting sessions. Roll calls were performed to make sure the council member was present. The data is updated when the duration of the meeting have been published.

The raw dataset features 10 variables and 6071 observations. @ tbl-raw-data shows the first 10 observations of the 10 variables: _id, Term, First Name, Last Name, Committee, MTG #, Session Date, Session Type, Session Start-End Time, and Present. ‘_id’ is the number associated with each observation. ‘Term’ shows which term the council members are a part of, which is 2022-2026 as this is the timeframe we are interested in. The ‘First name’ and ‘last name’ variables are the first and last names of each council member. ‘Committee’ is the committee the council member is a part of and in this paper, we are only looking at the City Council. ‘MTG #’ is the short form of meeting number. ‘Session Date’ refers to the day the session was held, in year-month-date format. ‘Session Type’ specifies whether the meeting took place during the Morning, Afternoon, or Evening. ‘Session Start-End Time’ gives the exact time the session was held. Finally, ‘Present’ marks whether the City Council member was present or absent.

## Cleaning Data

After importing the data into R (@ ) using the Open Data Toronto package (@ ), it is cleaned to keep only key variables of interest and make it easier for data analysis in Section 3. To clean the data and produce graphs with the data, the following packages in R (@ ) were used: tidyverse (@ ), dplyr (@ ), lubridate (@ ), knitr (@ ), and ggplot2 (@ ).

The key variables of interest are name, committee, session year, session type, and present. First, first and last names of City Council members are combined into 1 variable to make it easier to read in graphs. Then, the session date is coded to separate the year, month and date - keeping only the year. When cleaning the data, only the columns of interest were selected to be in the newly cleaned dataset. The cleaned dataset consists of 5 variables and 3061 observations as seen in @ tbl-cleaned-data. Tests are run on the cleaned dataset to ensure there are no empty rows & that the remaining observations consist of only City Council members’ data.

## Data Limitations

There were no similar datasets containing information about the attendance of Toronto City Council members on Open Data Toronto. If this paper were to be considered with different variables measuring the effectiveness of City Council members’ representation of the population then other datasets could be used. An example of a dataset that could have been an alternative is “City Council and Committees Meeting Schedule Reports” (@ ) which looks at the meeting schedule reports. The use of other datasets would have changed the research of this paper.

A limitation faced by this dataset is that some sessions are closed sessions. Closed sessions are those “closed to the public under sections 27-38 of the Council Procedures by law” (@ ). These closed sessions could be viewed as null observations potentially altering the results of data analysis. Another limitation of this dataset is the lack of information on the reason for absence. City Council members could have valid excuses such as illness or urgent emergencies. Without this information, the conclusions lack the reasoning behind the results.

# Results

## Toronto City Council member absence

```{r}
#| label: fig-session_type
#| fig-cap: Comparison of Attendance in Morning, Afternoon, vs Evening Session
#| echo: false

ggplot(cleaned_data, aes(x = session_type, fill = present)) +
  geom_bar(position = "dodge2") + 
  labs(
    title = "Present vs Absence in Different Session Types",
    x = "Session Type",
    y = "Count"
  )
```

## Toronto City Council absence over time since 2022

```{r}
#| label: fig-session_month
#| fig-cap: Comparison of Attendance over Different Months
#| echo: false

cleaned_data |> 
  ggplot(aes(x = session_month, fill = present)) +
  geom_bar(position = "dodge2") + 
  labs(
    title = "Present vs Absence Over Different Months",
    x = "Months",
    y = "Count"
  )
```

## Toronto City Council meeting type absence

```{r}
#| label: fig-session_year
#| fig-cap: Comparison of Attendance between 2022 to 2024
#| echo: false

cleaned_data |> 
  ggplot(aes(x = session_year, fill = present)) +
  geom_bar(position = "dodge") + 
  labs(
    title = "Present vs Absence Over Different Years",
    x = "Years",
    y = "Count"
  )
```

# Discussion

## Weaknesses and next steps

\newpage

\appendix

# Appendix {-}

## Raw Data

## Cleaned Data

\newpage

# References